SSA based global code motion of identical computations

### 2017-04-14 hoist.tex
register pressure model

\subsection{Reduce register pressure}
\label{hoist:reg-pressure}
Following example explains how code hoisting can actually reduce the register
pressure.  Consider the following example where the labels prefixed with 'P'
represent the position of instruction in a basic block (names prefixed with 'B').

\begin{verbatim}
B0: P0: b = 1
    goto B1

B1: P1: c = 2
    goto B2

B2: P2: if c is true then goto B3 else goto B4

B3: P3: a0 = b + c
    goto B5

B4: P4: a1 = b + c
    goto B5

B5: P5: d = phi {a0(B3), a1(B4)}

If we measure D(Px,Py) as total instruction
count in the path from the position of Px to Py

live-range(a0) = D(P0,P3) + D(P1,P3) + D(P3,P5)
               = 6 + 4 + 2 = 12
live-range(a1) = D(P0,P4) + D(P1,P4) + D(P4,P5)
               = 6 + 4 + 2 = 12

old-live-range = max(live-range(a0),live-range(a1))
               = 12

After hoisting a0 and a1 are removed and a copy
of a0 as a01 is placed in B2 just before P2.

live-range(a01) = D(P0,P2-1) + D(P1,P2-1) + D(P2 -1,P5)
                = 4 + 3 + 3 = 10
new-live-range = live-range(a01)
\end{verbatim}

If the new live-range is less than the old one it will be a good candidate for
hoisting. The live ranges can remain same as well if there is only one operand
on the right hand side, e.g., an assignment operation or, one of the operands is
not a register. That means hoisting upwards will decrease the live-range of its
use but increase the live-range of its definition.

In a special case where the instruction to be hoisted has the last use of its
operands then the code hoisting will always reduce the register pressure if it
has two register operands because the gain in live-range will be in the ratio of
2:1. Based on the above formulae we can also deduce that, as long as there is
one register operand in the right hand side with its last use, code hoisting
will either decrease or preserve the register pressure.

>>>>

\subsection{Reduce register pressure}
\label{hoist:reg-pressure}
Hoisting upwards will decrease the live-range of its use, if it is a last use (a
kill), but increase the live-range of its definition. Conversely, sinking will
decrease the live-range of the defined register but increase the live-range for
killed operands. If the live-range after code-motion is less than before it will
be moved.

Essentially, as long as there is one killed operand , code hoisting will either
decrease or preserve the register pressure. Similarly, code-sinking will either
decrease or preserve the register pressure as long as there is one operand
killed at most.

Following example explains how code motion of identical computations can reduce
the register pressure.  Consider the following example where the labels prefixed
with 'P' represent the position of instruction in a basic block (names prefixed
with 'B').

\begin{verbatim}
B0: b = m
    goto B1

B1: c = n
    goto B2

B2: if c is true then goto B3 else goto B4

B3: a0 = b<kill> + c<kill>

B4: a1 = b<kill> + c<kill>

After hoisting a0 and a1 are removed and a copy
of a0 as a01 is placed in B2 just before the branch.
\end{verbatim}

In this case, since `b' and `c' are killed in `a0' and `a1', hoisting them will
reduce the register pressure in B3 and B4 because two operands will be freed.
